\pagebreak
\subsection{Crotaphytidae --- Collared Lizard}
\begin{center}
\begin{longtabu} to \textwidth {| | p{3.5cm} | X | |}
	\hline
	Taxonomy/Ancestry &
	now considered a subfamily of iguanidae.
	\centeredgraphics{lacertila/crotaphytidae/tax}{0.5}
	 \\
	\hline
	Size & 
	8-15(20-38 cm) in length.
	\\
	\hline
	Color &
	collared --- bands of black around neck and shoulders. 
	 \\
	\hline
	Anatomy &
	\begin{itemize}[noitemsep]
		\item femoral pores
		\item small interparietal scale
		\item NO bony lines or projecting ridges
		\item long limbs and tails
		\item large head and powerful jaws
		\item live 5-8 years
		\item pleurodont teeth
	\end{itemize}	
	 \\
	\hline
	Dimorphism & 
	males have a blue-green body and light brown head. female's have a light brown head and body.
	\\
	\hline
	Behavior & 
	\begin{itemize}[noitemsep]
		\item quick runners; record = 16 mph (26 km/h)
		\item known for standing on hind legs and exhibit unusual form of saltatory bipedalism* to jump b/w boulders
		\item if 2 males are placed in the same cage, they will fight to the death
		\item undergo brumation
		\item emit squealing vocalizations when stressed, which is unknown among other iguanians
	\end{itemize}
	\\
	\hline
	Habitat & 
	desert-dwelling in arid, open regions.
	\\
	\hline
	Distribution & 
	native to Southwestern US and northern Mexico --- Missouri, Texas, Arizona, Kansas. full extent = Ozark Mountains to SoCal.
	\\
	\hline
	Feeding Ecology & 
	Crotaphytids are carnivores. Crotaphytus (collared lizard) species feed primarily on insect and small vertebrates. Species of Gambelia (the leopard lizards) are particularly noted for feeding on other reptiles, especially smaller iguanian lizards.
	\\
	\hline
	Reproductive Biology & 
	In a few species (e.g. Gambelia wislizeni, Crotophytus collaris), the females' orange spots brighten during shortly before oviposition, and fade between clutches. In at least one Crotaphytus species, males have been observed licking females during courtship.
	\\
	\hline
	Ecological Role &
	predators: other lizards, birds (eg roadrunners), coyotes, cats, carnivorous mammals.
	\\
	\hline
	Conservation Status & 
	not threatened or endangered. common collared lizard = Oklahoma state reptile. known as ``mountain boomer," possibly b/c settlers mistook sound of wind for lizard calls.
	\\
	\hline
\end{longtabu}
\scalegraphics{lacertila/crotaphytidae/1}{0.85}
\scalegraphics{lacertila/crotaphytidae/2}{0.25}
\end{center}